---
title: "Single species, single site models"
author: "Richard Erickson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single species, single site models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**Warning:** Running this vignette the first time will take several minutes for the different models to compile.


I created these models as building blocks with Stan to help myself understand the Stan code.
These simple models are building blocks for more complex models.
I am sharing them to help other people understand Stan models.
The models do not include coefficients, but could be easily adapted to include these by a person familiar with Stan code.


```{r}
#| eval: true
#| echo: false
#| warning: false
#| message: false
#| error: false
library(occstanhm)
library(tidyverse)
```

## psi only model

The first model only fits a $\psi$-level model.
The model has multiple revisits (`n_revisits`) with binary response (`z`) to a single site that allows $\psi$ to be estimated on the logit scale as `logit_psi`.
Mathematically, this model would be:

$$
Z_{i} \sim \textrm{Bernoulli}(\psi),~\textrm{where}~i \in N_{\textrm{revisits}}.
$$
The model would be the same as a logistic regression, but is coded to allow other levels to be added to the model.
Some key points about the model's Stan code:  

* The model loops over each revisit from a single site.
* The `logit_psi` is converted to the log-scale to make it additive.
* 1 - $\psi$  on the logit scale is converted to the log-scale for the previous reasons and to increase the speed of computations.

And, the model as coded in Stan:

```{r comment=''}
#| echo: false
system.file("./stan_models/tutorial/occ_1lvl_1site_1spp.stan",
            package = "occstanhm") |>
  readLines() |>
  cat(sep = "\n")
```

This model may be fit using this package in R.
The simulation includes a large number to ensure consistent results with the simulation and recovery of parameter estimates.
Notice how the simulated values for `psi` is what the model recovers on the logit scale.

```{r}
library(occstanhm)
library(tidyverse)
psi <- 0.25
n_revisits <- 400
z <- rbinom(n = n_revisits, size = 1, prob = psi)
n_chains <- 4
n_parallel_chains <- 4
n_threads_per_chain <- 4
n_refresh <- 0
n_warmup <- 2000
n_sample <- 2000

stan_data <-
  list(z = z,
       n_revisits = n_revisits)

psi_fit <-
  occ_1lvl_1site_1spp(stan_data = stan_data,
                      n_chains = n_chains,
                      n_threads_per_chain = n_threads_per_chain,
                      n_warmup = n_warmup,
                      n_sample = n_sample,
                      n_refresh = n_refresh)

## examine at model estiamted value
psi_fit

## examine simulated value from the model
mean(z) |> qlogis() |> round(2)
```

## psi and theta  model

Next, the model can be expanded to include both a $\psi$ and $\theta$-level.
This model estimates both site occupancy ($\psi$; `logit_psi` on the logit scale) and a sample capture probability ($\theta$; `logit_theta` on the logit scale).
That includes the total number of samples from the site (`n_total_samples`),
the number of revisits to the site (`n_revisits`),
the number of samples per revisit (`n_samples_per_revisit`),
the observed occupancy or nonoccupancy at a site during a revisit (`z_obs`),
the capture or non-capture in a sample (`a`),
and a subsetting for each revisits start (`unit_start`) and revisits stop (`unit_stop`) in the `a` vector.
This last parameter is necessary for Stan's discrete nature for latent variables.


The equations are similar to the previous model, but also include the sample capture probability, which is conditional (the $|$ symbol) upon the site being occupied:

\begin{align}
Z_{i} &\sim \textrm{Bernoulli}(\psi),~\textrm{where}~i \in N_{\textrm{revisits}}~\textrm{and} \\
A_{i,j} | z_{i} &\sim \textrm{Bernoulli}(z_{i} \theta),~\textrm{where}~j \in N_{\textrm{samples per revisits}}.
\end{align}

For the Stan code, the model expands upon the previous model and includes the following extras:

* The code loads a custom function, `detection_lpmf`.
* The custom `detection_lpmf` function (contents not shown) contains a second loop over samples within each site.
* These nesting samples exist to allow the addition of extra layers to the model.
* The observed non-capture is either the probability the species was not there, or, the probability the species was there, but missed.

Now, the Stan code:

```{r comment=''}
#| echo: false
system.file("./stan_models/tutorial/occ_2lvl_1site_1spp.stan",
            package = "occstanhm") |>
  readLines() |>
  cat(sep = "\n")
```

This model may be fit using this package in R.
Notice how the simulated values for `psi` and `theta` are close to what the model recovers on the logit scale:

```{r}
library(occstanhm)
library(tidyverse)
psi <- 0.75
n_revisits <- 30
n_samples <- rpois(n_revisits, lambda = 10) + 1L
# +1 is to ensure at least 1 sample
n_samples
theta_sim <- 0.6
dat <-
  tibble(unit = "one",
         revisit = seq(1, n_revisits),
         z = rbinom(n = n_revisits, size = 1, prob = psi)) |>
  full_join(tibble(unit = "one",
                   revisit = rep(seq(1, n_revisits), times = n_samples),
                   theta =  theta_sim),
            by = c("unit", "revisit")) |>
  mutate(a = z * rbinom(n(), 1, theta)) |>
  group_by(unit, revisit) |>
  mutate(z_obs = ifelse(sum(a) > 0, 1, 0)) |>
  ungroup() |>
  rowid_to_column("index")
dat

sample_summary <-
  dat |>
  group_by(unit, revisit) |>
  summarize(n_samples_per_revisit = n(),
            theta_obs = mean(a),
            z_obs = ifelse(sum(a) > 0, 1, 0),
            unit_start = min(dplyr::cur_group_rows()),
            unit_stop = max(dplyr::cur_group_rows()),
            .groups = "drop")

stan_data <-
  list(a = dat |> pull(a),
       z_obs = sample_summary |> pull(z_obs),
       n_revisits = sample_summary |> nrow(),
       n_samples_per_revisit = sample_summary |> pull(n_samples_per_revisit),
       unit_start = sample_summary |> pull(unit_start),
       unit_stop = sample_summary |> pull(unit_stop),
       n_total_samples = dat |> nrow())

n_chains <- 4
n_parallel_chains <- 4
n_threads_per_chain <- 4
n_refresh <- 0
n_warmup <- 2000
n_sample <- 2000

fit <- occ_2lvl_1site_1spp(stan_data = stan_data,
                           n_chains = n_chains,
                           n_parallel_chains = n_parallel_chains,
                           n_threads_per_chain = n_threads_per_chain,
                           n_refresh = n_refresh,
                           n_warmup = n_warmup,
                           n_sample = n_sample)

fit

## approximate simulated theta on logit scale
dat |>
  filter(z_obs > 0) |>
  summarize(mean_a = mean(a)) |>
  pull(mean_a) |>
  qlogis()

## approximate simulated psi on logit scale
sample_summary |>
  summarize(obs_psi = mean(z_obs)) |>
  mutate(logit_psi = qlogis(obs_psi))
```

**Tip:** Use the R code to debug and understand the subsetting/indexing loops. You can write R code to be very similar to Stan.
For example, with the above code, you might use `stan_data$unit_start` and `stan_data$unit_stop` with a for-loop to check your indexing with real data.

## psi, theta, and p  model

The last is a single site, single species, three-level occupancy model.
This model includes the site's occupancy, $psi$, the sample capture probability, $\theta$, and the detection probability within each subsample, $p$.  
The subsample observations ($y$) come from a binomial distribution with $k$ subsamples per sample.
This level of the model allows us to avoid a for loop (at the cost of including individual-level coefficients for each subsample).

```{=tex}
\begin{align}
Z_{i} &\sim \textrm{Bernoulli}(\psi),~\textrm{where}~i \in N_{\textrm{revisits}}, \\
A_{i,j} | z_{i} &\sim \textrm{Bernoulli}(z_{i} \theta),~\textrm{where}~j \in N_{\textrm{samples per revisits}}, ~\textrm{and}\\
Y_{i,j} | a_{i,j} &\sim \textrm{Binomial}(a_{i,j} p | k_{i,j}).
\end{align}
```

With the Stan code, the results are similar to the previous model.
Because we can use a binomial distribution, the model does not require a third loop:


```{r comment=''}
#| echo: false
system.file("./stan_models/tutorial/occ_3lvl_1site_1spp.stan",
            package = "occstanhm") |>
  readLines() |>
  cat(sep = "\n")
```

This model may be fit using this package in R.
Notice how the simulated values for `psi`, `theta`, and `p` are close to what the model recovers on the logit scale.

```{r}
library(tidyverse)

psi_sim <- 0.75
n_revisits <- 40
n_samples <- rpois(n_revisits, lambda = 10) + 1L
n_samples
theta_sim <- 0.6
theta_sim
n_subsamples_in <- 10
p_sim <- 0.25

dat <-
  tibble(unit = "one",
         revisit = seq(1, n_revisits),
         z = rbinom(n = n_revisits, size = 1, prob = psi_sim),
         n_subsamples = n_subsamples_in) |>
  full_join(tibble(unit = "one",
                   revisit = rep(seq(1, n_revisits), times = n_samples),
                   theta =  theta_sim,
                   p = p_sim),
            by = c("unit", "revisit")) |>
  mutate(a = z * rbinom(n(), 1, theta),
         y = a * rbinom(n(), n_subsamples, p)) |>
  group_by(unit, revisit) |>
  mutate(a_obs = ifelse(a > 0, 1, 0),
         z_obs = ifelse(sum(y) > 0, 1, 0)) |>
  ungroup() |>
  rowid_to_column("index")

dat

dat |>
  filter(z != z_obs)

dat |>
  filter(a > 0)

dat |>
  filter(a != a_obs)

sample_summary <-
  dat |>
  group_by(unit, revisit) |>
  summarize(n_samples_per_revisit = n(),
            p_obs = mean(y / n_subsamples),
            theta_obs = mean(a_obs),
            z_obs = ifelse(sum(a_obs) > 0, 1, 0),
            unit_start = min(dplyr::cur_group_rows()),
            unit_stop = max(dplyr::cur_group_rows()),
            .groups = "drop")

sample_summary

stan_data <-
  list(y = dat |> pull(y),
       a_obs = dat |> pull(a_obs),
       k_subsamples = dat |> pull(n_subsamples),
       z_obs = sample_summary |> pull(z_obs),
       n_revisits = sample_summary |> nrow(),
       n_samples_per_revisit = sample_summary |> pull(n_samples_per_revisit),
       unit_start = sample_summary |> pull(unit_start),
       unit_stop = sample_summary |> pull(unit_stop),
       n_total_samples = dat |> nrow())

n_chains <- 4
n_parallel_chains <- 4
n_threads_per_chain <- 4
n_refresh <- 0
n_warmup <- 2000
n_sample <- 2000

fit <- occ_3lvl_1site_1spp(stan_data,
                           n_chains = n_chains,
                           n_parallel_chains = n_parallel_chains,
                           n_threads_per_chain = n_threads_per_chain,
                           n_refresh = n_refresh,
                           n_warmup = n_warmup,
                           n_sample = n_sample)

fit_summary <-
  fit$summary()

fit_summary |>
  filter(variable %in% c("logit_psi", "logit_theta", "logit_p")) |>
  select(variable, mean, median) |>
  mutate(p_mean = plogis(median))

dat
psi_sim
theta_sim
p_sim

sample_summary |>
  summarize(obs_psi = mean(z_obs)) |>
  mutate(logit_psi = qlogis(obs_psi))
```
