---
title: "Single species, multiple site models"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single species, multiple site models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#| message: false
#| error: false
library(occstanhm)
library(tidyverse)
```

**Warning:** Running this vignette the first time will take several minutes for the different models to compile.

I (RAE) created these models as building blocks with Stan to help myself understand the Stan code.
These simple models are building blocks for more complex models.
I am sharing them to help other people understand Stan models.
The models do not include coefficients, but could be easily adapted to include these by a person familiar with Stan code.

This vignette covers models with single species from multiple locations.
These locations may be replicated across space or time and are called _units_ in this package, although they could also be called _sites_ or similar terms.


## Multi-psi model

The first model estimates multiple values for $\psi_j$ (`psi`, or `logit_psi` on the logit scale in Stan) from $N_\textrm{units}$ different units (`n_units`), $j$ (`unit_idx`).
Each unit may have a unique number of revisits that are stored in vector, $N_\textrm{revisits per unit}$, `n_revisits_per_unit`:

$$
Z_{i,j} \sim \textrm{Bernoulli}(\psi_j),~\textrm{where}~i \in N_{\textrm{revisits per unit},~j}.
$$

The model has multiple binary revisits (`z`) that are sliced using a starting index vector (`unit_start`) and a stopping index vector (`unit_stop`) for a single site that allows each $\psi_j$ to be estimated on the logit scale as a vector `logit_psi`.
In Stan, this is may be coded as follows:


```{r comment=''}
#| echo: false
system.file("stan_models/tutorial/occ_1lvl_multisite_1spp.stan",
            package = "occstanhm") |>
  readLines() |>
  cat(sep = "\n")
```

Data may be simulated and then the model fit with R as well:

```{r}
library(occstanhm)
library(tidyverse)
n_units_sim <- 10 # max is 26 due to number of letters
psi_vec_in <- rbeta(n_units_sim, 3, 3)
hist(psi_vec_in)
n_units_in <- length(psi_vec_in)

n_revisits_per_unit_in <- rpois(n_units_sim, 35)

dat <-
  tibble(unit = rep(letters[seq(1, n_units_in)],
                    times = n_revisits_per_unit_in),
         psi_vec = rep(psi_vec_in, times = n_revisits_per_unit_in)) |>
  mutate(z_obs =  rbinom(n(), size = 1, prob = psi_vec)) |>
  rowid_to_column("index") |>
  ungroup()

## Need to make sure arrange by unit first
## then within unit
unit_summary <-
  dat |>
  group_by(unit) |>
  summarize(psi_obs = mean(z_obs),
            unit_start = min(index),
            unit_stop = max(index),
            n_revisits_per_unit = n()) |>
  mutate(logit_psi = qlogis(psi_obs))

unit_summary

stan_data <-
  list(
    z_obs = dat |> pull(z_obs),
    n_revisits = dat |> nrow(),
    n_revisits_per_unit = unit_summary |> pull(n_revisits_per_unit),
    n_units = unit_summary |> nrow(),
    unit_start = unit_summary |> pull(unit_start),
    unit_stop = unit_summary |> pull(unit_stop),
    grainsize = 1
  )

n_chains <- 4
n_parallel_chains <- 4
n_threads_per_chain <- 4
n_refresh <- 0
n_warmup <- 2000
n_sample <- 2000

fit <- occ_1lvl_multisite_1spp(stan_data = stan_data,
                               n_chains = n_chains,
                               n_parallel_chains = n_parallel_chains,
                               n_threads_per_chain = n_threads_per_chain,
                               n_refresh = n_refresh,
                               n_warmup = n_warmup,
                               n_sample = n_sample)

fit
unit_summary |> pull(psi_obs) |> qlogis() |> round(2)
unit_summary |> pull(psi_obs) |> round(2)
```

## occupancy_2

A second model estimates multiple values for $\psi_j$ (`psi`, or `logit_psi` on the logit scale in Stan) and $\theta$ (`theta` or `logit_theta` on the logit scale in Stan) from $N_\textrm{units}$ different units (`n_units`), $j$ (`unit_idx`).
Each unit may have a unique number of revisits that are stored in vector, $N_\textrm{revisits per unit}$, `n_revisits_per_unit`.
Likewise, each unit may have a unique number of samples per revisit that are stored in an object (a matrix in math syntax and a vector in Stan), $N_{\textrm{samples per revisit},~j, ~k}$ (`n_samples_per_revisit`): 

```{=tex}
\begin{align}
Z_{i,j} &\sim \textrm{Bernoulli}(\psi_j),~\textrm{where}~i \in N_{\textrm{revisits per unit},~j}~\textrm{and}\\
A_{i,j,k}|z_{i,j}  &\sim \textrm{Bernoulli}(z_{i,j} \theta_{j})~\textrm{where}~i \in N_{\textrm{samples per revisit},~j,~k}\\
\end{align}
```

Adapting this code to Stan builds upon the previous model by including a second set of subsetting.
With this syntax, `unit_revisit_start` and `unit_revisit_stop` slice revisit-level data by each unit and `unit_dat_start` and `unit_dat_stop` slice the observed ($a$-level) data by each unit.
`revisit_start_in_unit` and `revisit_stop_in_unit` slice each revisit within a unit for each revisit.

**Warning:** Nested for-loops are confusing and tricky in Stan.
They exist in this model because of Stan's syntax for the marginalization of discrete latent variables and the parallel `reduce_sum()` function.

First, these are combined for slice data by unit (for the first for-loop in Stan).
Second, these are used for each revisit (for the second for-loop in Stan).
Third, a loop goes over each sample (for the third for-loop in Stan):
These functions are then coded in Stan:

```{r comment=''}
#| echo: false
system.file("stan_models/tutorial/occ_2lvl_multisite_1spp.stan",
            package = "occstanhm") |>
  readLines() |>
  cat(sep = "\n")
```

This Stan model may then be fit in R after simulating data:

```{r}
library(tidyverse)
n_units_sim <- 5 # max is 26 due to number of letters
psi_vec_in <- rbeta(n_units_sim, 3, 3)
theta_vec_in <- rbeta(n_units_sim, 3, 3)

hist(psi_vec_in)
hist(theta_vec_in)

n_units_in <- length(psi_vec_in)

n_revisits_per_unit_in <- rpois(n_units_sim, 30)
hist(n_revisits_per_unit_in)

n_samples_per_revisits <- rpois(sum(n_revisits_per_unit_in), 30) + 1L
hist(n_samples_per_revisits)

## unit and revisit data
dat_revisit <-
  tibble(unit = rep(letters[seq(1, n_units_in)],
                    times = n_revisits_per_unit_in),
         psi_vec = rep(psi_vec_in, times = n_revisits_per_unit_in),
         theta_vec = rep(theta_vec_in, times = n_revisits_per_unit_in)) |>
  mutate(z =  rbinom(n(), size = 1, prob = psi_vec)) |>
  rowid_to_column("index") |>
  group_by(unit) |>
  mutate(revisit_id = dplyr::cur_group_rows() -
           min(dplyr::cur_group_rows()) + 1L) |>
  ungroup() |>
  select(unit, revisit_id, psi_vec, theta_vec, z)

# sample
dat_sample <-
  tibble(
    unit = rep(dat_revisit |> pull(unit),
               times = n_samples_per_revisits),
    revisit_id = rep(dat_revisit |> pull(revisit_id),
                     times = n_samples_per_revisits),
  ) |>
  rowid_to_column("index") |>
  group_by(unit) |>
  mutate(sample_id = dplyr::cur_group_rows() -
           min(dplyr::cur_group_rows()) + 1L) |>
  select(-index) |>
  ungroup()

# merge together
dat <-
  dat_revisit |>
  full_join(dat_sample, by = c("unit", "revisit_id")) |>
  rowid_to_column("index") |>
  mutate(a = z * rbinom(n(), 1, theta_vec)) |>
  group_by(unit, revisit_id) |>
  mutate(z_obs = ifelse(sum(z) > 0, 1, 0)) |>
  ungroup()

## Need to make sure arrange by unit first
## then within unit
unit_summary <-
  dat |>
  group_by(unit) |>
  summarize(psi_obs = mean(z_obs),
            unit_dat_start = min(index),
            unit_dat_stop = max(index),
            n_revisits_per_unit = n()) |>
  mutate(logit_psi = qlogis(psi_obs))

unit_summary

revisit_summary <-
  dat |>
  group_by(unit, revisit_id) |>
  summarize(n_samples_per_revisit = n(),
            revisit_start = min(index),
            revisit_stop = max(index),
            z_obs = mean(z_obs),
            .groups = "drop") |>
  group_by(unit) |>
  mutate(revisit_start_in_unit = revisit_start - min(revisit_start) + 1L,
         revisit_stop_in_unit = revisit_stop - min(revisit_start) + 1L)

unit_revisit_summary <-
  revisit_summary |>
  rowid_to_column("index") |>
  group_by(unit) |>
  summarize(unit_revisit_start = min(index),
            unit_revisit_stop = max(index),
            n_samples_per_unit = n())
stan_data <-
  list(
    a = dat |> pull(a),
    unit_dat_start = unit_summary |> pull(unit_dat_start),
    unit_dat_stop = unit_summary |> pull(unit_dat_stop),
    revisit_start_in_unit = revisit_summary |> pull(revisit_start_in_unit),
    revisit_stop_in_unit = revisit_summary |> pull(revisit_stop_in_unit),
    n_total_samples = dat |> nrow(),
    z_obs = revisit_summary |> pull(z_obs),
    n_samples_per_revisit = revisit_summary |> pull(n_samples_per_revisit),
    n_total_revisits = revisit_summary |> nrow(),
    unit_revisit_start = unit_revisit_summary |> pull(unit_revisit_start),
    unit_revisit_stop = unit_revisit_summary |> pull(unit_revisit_stop),
    n_revisits_per_unit = unit_revisit_summary |> pull(n_samples_per_unit),
    n_samples_per_unit = unit_summary |> pull(n_revisits_per_unit),
    n_units = unit_summary |> nrow(),
    grainsize = 1
  )

n_chains <- 4
n_parallel_chains <- 4
n_threads_per_chain <- 4
n_refresh <- 0
n_warmup <- 1000
n_sample <- 1000

fit <- occ_2lvl_multisite_1spp(stan_data,
                               n_chains = n_chains,
                               n_parallel_chains = n_parallel_chains,
                               n_threads_per_chain = n_threads_per_chain,
                               n_refresh = n_refresh,
                               n_warmup = n_warmup,
                               n_sample = n_sample)

fit$summary() |>
  filter(grepl("logit_psi|logit_theta", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean))

unit_summary |> pull(psi_obs) |> qlogis() |> round(2)
unit_summary |> pull(psi_obs) |> round(2)

theta_vec_in |> round(2)
```
