---
title: "Correlated error structure models"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Correlated error structure models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#| message: false
#| error: false
library(occstanhm)
library(tidyverse)
```

The models in this vignette include a correlated error structure for species across locations.

**Warning:** The examples in this vignette are designed to show increasing levels of computational complexity.
The models do not include generalized coefficients, and, in fact all coefficients are coded to match the number of units and species.
The generalized versions of these models allow for greater flexibility.
These models are included in the package as learning artifacts more so than production models.
One advantage of these example's lack of generalization is a quicker model because the `logit_` parameters are the same length as the number of unit-species combinations whereas the generalized model's parameter are the same length as the number of total observations. 
This compromise shows the trade offs of "hard coding" models for custom applications compared to generalized versions with respect to application to new problems versus performance speed.

## 1-level model

A one-level model is not included in this package.
An example appears in the `fishStan` package with the hierarchical logistic regression (Erickson et al. 2020; 2022).


## 2-level model

This is a two-level model that estimates an intercept for each species at each location for both site occupancy ($\psi$) and site detection ($\theta$).
Specifically, the model estimates multiple $\psi$ values for multiple species, $\psi_{j,s}$ (`psi`, or `logit_psi` on the logit scale in Stan) and multiple $\theta$ values for multiple species $\theta_{j,s}$ (`theta`, or `logit_theta` on the logit scale in Stan) from  $N_\textrm{units}$ different units (`n_units`), $j$ (`unit_idx`), and $N_\textrm{species}$ different species (`n_species`).
The indices are combined in Stan as well to be $N_\textrm{units, species}$.
This indexing is needed because of how Stan vectorizes and slices data.
Each unit-species combination may have a unique number of revisits that are stored in the vector, $N_\textrm{revisits per unit}$, `n_revisits_per_unit`:

```{=tex}
\begin{align}
Z_{i,~j,~s} &\sim \textrm{Bernoulli}(\textrm{logit}(\psi_{j, s})),~\textrm{where}~i \in N_{\textrm{revisits per unit},~j},~j\in N_\textrm{number units},~\textrm{and}~s \in N_\textrm{species},~\textrm{and} \\
A_{i,~j,~k~s}|z_{i,~j,~s} &\sim \textrm{Bernoulli}(\textrm{logit}(z_{i,~j,~s}\theta_{j, s}))~\textrm{where}~k \in N_{\textrm{samples per revisit}}.
\end{align}
```

The model includes a predictor matrix for $\textrm{logit}(\psi_{j, s})$ (`logit_psi`) of coefficients $\boldsymbol{\beta}$.
The model includes a predictor matrix for $\textrm{logit}(\theta_{j, s})$ (`logit_theta`) of coefficients $\boldsymbol{\alpha}$.
Each unit has its own row in a predictor vector $\boldsymbol{\alpha}$ and each species is a column in $\boldsymbol{\alpha}$.
A vector of length $N_{\textrm{unit, species}, jj_\psi}$, expands $\boldsymbol{\beta}$ to match the vector $\textrm{logit}(\psi_{j, s})$.
A vector of length $N_{\textrm{total observations}, jj_\textrm{p}}$, expands $\delta$ to match the vector $\textrm{logit}(\theta_{j, s})$.
Lastly, a predictor matrices includes binary membership coding for each species, $\textbf{X}$ and $\textbf{W}$, at the two levels:

```{=tex}
\begin{align}
\textrm{logit}(\psi_{j, s}) &= \textbf{X} \boldsymbol{\beta}[jj_\psi]~\textrm{and}\\
\textrm{logit}(\theta_{j, s}) &= \textbf{W} \boldsymbol{\alpha}[jj_\theta].
\end{align}
```

These two parameters are assigned priors, $\sigma$ so that

```{=tex}
\begin{align}
\textrm{logit}(\psi_{j, s}) &\sim \textrm{Normal}(\textbf{X} \boldsymbol{\beta}[jj_\psi], \sigma_\psi)~\textrm{and}\\
\textrm{logit}(\theta_{j, s}) &\sim \textrm{Normal}(\textbf{W} \boldsymbol{\alpha}[jj_\theta], \sigma_\theta).
\end{align}
```

Each prior is also assigned its own prior as well to help with numerical convergence:

```{=tex}
\begin{align}
\sigma_\psi &\sim \textrm{Half-normal}(0, \sigma_{\psi~\textrm{prior}})~\textrm{and} \\
\sigma_\theta &\sim \textrm{Half-normal}(0, \sigma_{\theta~\textrm{prior}}).
\end{align}
```

The coefficients from each site are drawn from a multivariate distribution with a mean vector and covariance matrix with its own predictors:

```{=tex}
\begin{align}
\boldsymbol{\beta} &\sim \textrm{multivariate normal}(\textbf{X}^\star \boldsymbol{\beta}^\star, \boldsymbol{\Sigma}^\psi),~\textrm{and} \\
\boldsymbol{\alpha} &\sim \textrm{multivariate normal}(\textbf{W}^\star \boldsymbol{\alpha}^\star, \boldsymbol{\Sigma}^\theta). \\
\end{align}
```

The variance/covariance matrices are proportional to the correlation matrices among coefficients (in this case, the species-specific occupancy probabilities) $\boldsymbol{\Omega}$ so that:

```{=tex}
\begin{align}
\boldsymbol{\Sigma}^\psi &\propto \boldsymbol{\Omega}^\psi~\textrm{and}\\
\boldsymbol{\Sigma}^\theta &\propto \boldsymbol{\Omega}^\theta.\\
\end{align}
```

The derivation and numerical implantation of this relationship are presented in[\S1.13 of the Stan User Guide](https://mc-stan.org/docs/stan-users-guide/regression.html#multivariate-hierarchical-priors.section).
This model may be implemented in Stan.
The model requires a custom Cholesky factor function (located in the `scale_cholesky_factor.stan` file) because the the default [Stan method does not scale well past 5-10 coefficients](https://discourse.mc-stan.org/t/nan-results-with-lkj-corr-cholesky/16832), and, instead requires the [onion method](https://discourse.mc-stan.org/t/using-the-onion-method-in-occupancy-model-hierarchical-model/24901).

The model inputs are as follows:

* the model requires the number of unit-species combinations (`n_unit_species`)
* the total number of samples (`n_total_samples`),
* the number of units (`n_units`; **WARNING**: this is included as part of a simplifying assumption about input and predictor dimensions and more complex models do not require this parameter),
* a vector with the number of revisits per unit species combination (`n_revisits_per_unit_species`),
* the total number of all revisit species combinations (`n_total_revisits_species`),
* a vector with the number of samples for each revisit species combination (`n_samples_per_revisit_species`),
* a vector with the binary observation or nonobservation of a species at a revisit (`z_obs`),
* and a vector with the binary detection or nondetection within a sample (`a`).
* The model also requires double indexing (that is, indexing within indexing due to Stan's challenging syntax as a language that does not easily handle discrete latent variables and also the use of the `reduce_sum()` function):

  - a indexing vector for the start (`unit_species_revisit_start`) and stop (`unit_species_revisit_stop`) of unit species combinations at the revisit level data (such as `z_obs`),
  - an indexing for the start (`unit_species_dat_start`) and stop (`unit_species_dat_stop`) of unit species combinations at the data level (such as `a`), and
  - an indexing of the start (`revisit_start_in_unit`) and stop (`revisit_stop_in_unit`) of each revisit within a unit. The double indexing uses one index to subset within another index as part of the code.

* the number of columns in $\beta$ (`m_beta`) and $\alpha$ (`m_alpha`; **NOTE:** The generalized version of the model include row vectors, but this model uses other variables like `n_units`)
* an expansion vector for $\psi$-level predictors (`jj_psi`)
* a predictor matrix $\textbf{X}$ for the $\psi$-level (`x_psi`; **NOTE::** many computer variables in the model use longer names to make them easier to read and find using tools like "find-and-replace" like this variable name)
* a hyperparameter column number, `m_beta_star`
* a hyperparameter predictor matrix $\textbf{X}^\star$ (`x_psi_star`)
* an expansion vector for $\theta$-level predictors (`jj_theta`)
* a predictor matrix $\textbf{W}$ for the $\theta$-level (`w_theta`)
* a hyperparameter column number, `m_alpha_star`
* a hyperparameter predictor matrix $\textbf{W}^\star$ (`w_theta_star`)
* a prior matrix, `beta_psi_star_prior`
* a prior `beta_psi_star_sd_prior`
* a prior for the Cholesky matrix `eta_psi`
* a prior matrix `alpha_theta_star_prior`
* a prior `alpha_theta_star_sd_prior`
* a prior for the Cholesky matrix `eta_theta`
* a setting for the `reduce_sum()` function, `grainsize`

The model is then coded with Stan:

```{r comment=''}
#| echo: false
system.file("stan_models/tutorial/occ_2lvl_multispp_corr.stan",
            package = "occstanhm") |>
  readLines() |>
  cat(sep = "\n")
```

The model may be run in R.
**Note:** The model and simulated data have been pre-run and included in the package to speed computational time.
First, the data is simulated:

```{r}
#| eval: false
library(occstanhm)
library(tidyverse)

set.seed(123)
n_units_sim <- 2 # max is 26 due to number of letters
n_species_sim <- 4

n_revisits_per_site_in <- rpois(n_units_sim, 30)
hist(n_revisits_per_site_in)

n_samples_per_revisits <- rpois(sum(n_revisits_per_site_in), 20) + 1L
hist(n_samples_per_revisits, breaks = seq(0, 50, by = 1))

## species unit
dat_species <-
  expand_grid(unit = rep(letters[seq(1, n_units_sim)]),
              species = paste("species", seq(1, n_species_sim))) |>
  mutate(psi_vec = rbeta(n(), 3, 3),
         theta_vec = rbeta(n(), 3, 3))

dat_species |>
  print(n = Inf)

## unit and revisit data
dat_revisit <-
  tibble(unit = rep(letters[seq(1, n_units_sim)],
                    times = n_revisits_per_site_in)) |>
  rowid_to_column("index") |>
  group_by(unit) |>
  mutate(revisit_id = dplyr::cur_group_rows() -
           min(dplyr::cur_group_rows()) + 1L) |>
  ungroup() |>
  select(unit, revisit_id)

dat_revisit |>
  print(n = Inf)

# sample
dat_sample <-
  tibble(
    unit = rep(dat_revisit |> pull(unit),
               times = n_samples_per_revisits),
    revisit_id = rep(dat_revisit |> pull(revisit_id),
                     times = n_samples_per_revisits),
  ) |>
  rowid_to_column("index") |>
  group_by(unit) |>
  mutate(sample_id = dplyr::cur_group_rows() -
           min(dplyr::cur_group_rows()) + 1L) |>
  select(-index) |>
  ungroup()

# merge together
dat <-
  dat_species |>
  full_join(dat_revisit, by = "unit",
            relationship = "many-to-many") |>
  mutate(z = rbinom(n(), 1, psi_vec)) |>
  full_join(dat_sample, by = c("unit", "revisit_id"),
            relationship = "many-to-many") |>
  arrange(unit, species) |>
  rowid_to_column("index") |>
  mutate(a = z * rbinom(n(), 1, theta_vec)) |>
  ungroup() |>
  group_by(unit, revisit_id, species) |>
  mutate(z_obs = ifelse(sum(a) > 0, 1, 0)) |>
  ungroup()

unit_species_summary <-
  dat |>
  group_by(unit, species) |>
  summarize(psi_obs = mean(z_obs),
            unit_species_dat_start = min(index),
            unit_species_dat_stop = max(index),
            n_revisits_per_site = n(),
            .groups = "keep") |>
  mutate(logit_psi = qlogis(psi_obs))

unit_species_summary |>
  print(n = Inf)

revisit_species_summary <-
  dat |>
  group_by(unit, species, revisit_id) |>
  summarize(n_samples_per_revisit = n(),
            revisit_start = min(index),
            revisit_stop = max(index),
            z_obs = mean(z_obs),
            .groups = "drop") |>
  group_by(unit, species) |>
  mutate(revisit_start_in_unit_species = revisit_start -
           min(revisit_start) + 1L,
         revisit_stop_in_unit_species = revisit_stop -
           min(revisit_start) + 1L)

unit_species_revisit_summary <-
  revisit_species_summary |>
  rowid_to_column("index") |>
  group_by(unit, species) |>
  summarize(unit_species_revisit_start = min(index),
            unit_species_revisit_stop = max(index),
            n_revisits_per_unit_species = n(),
            psi_obs = mean(z_obs),
            .groups = "drop") |>
  mutate(unit_id = as.integer(factor(unit)))

```

Then the data is formatted for Stan:
```{r}
#| eval: false
## Setup parameters for model

# psi-level
x_psi <- model.matrix(~ species - 1, unit_species_summary)
x_psi |> head()
x_psi |> dim()

unit_species_summary_hyper <-
  unit_species_summary |>
  group_by(unit) |>
  summarize(n = n(),
            hyper_psi = mean(psi_obs),
            hyper_logit = mean(logit_psi))

x_psi_star <-
  model.matrix(~ 1, unit_species_summary_hyper)

beta_psi_star_prior <-
  matrix(0,
         nrow = x_psi_star |> ncol(),
         ncol = x_psi |> ncol())

# theta-level
w_theta <- model.matrix(~ species - 1, unit_species_summary)
w_theta
w_theta |> head()
w_theta |> dim()

w_theta_star <-
  model.matrix(~ 1, unit_species_summary_hyper)

alpha_theta_star_prior <-
  matrix(0,
         nrow = w_theta_star |> ncol(),
         ncol = w_theta |> ncol())

## Save data in list for stan
stan_data <-
  # observation and latent observation data
  list(
    a = dat |> pull(a),
    z_obs = revisit_species_summary |> pull(z_obs),
    # indexing
    unit_species_dat_start = unit_species_summary |>
      pull(unit_species_dat_start),
    unit_species_dat_stop = unit_species_summary |>
      pull(unit_species_dat_stop),
    revisit_start_in_unit = revisit_species_summary |>
      pull(revisit_start_in_unit_species),
    revisit_stop_in_unit = revisit_species_summary |>
      pull(revisit_stop_in_unit_species),
    unit_species_revisit_start = unit_species_revisit_summary |>
      pull(unit_species_revisit_start),
    unit_species_revisit_stop = unit_species_revisit_summary |>
      pull(unit_species_revisit_stop),
    # all different dimensions, such as "n"
    n_units = dat |> distinct(unit) |> nrow(),
    n_units_species = unit_species_summary |> nrow(),
    n_total_samples = dat |> nrow(),
    n_revisits_per_unit_species = unit_species_revisit_summary |>
      pull(n_revisits_per_unit_species),
    n_samples_per_revisit_species = revisit_species_summary |>
      pull(n_samples_per_revisit),
    n_total_revisits_species = revisit_species_summary |> nrow(),

    # psi-level inputs
    x_psi = x_psi,
    m_beta = x_psi |> ncol(),
    x_psi_star = x_psi_star,
    m_beta_star = x_psi_star |> ncol(),
    beta_psi_star_prior = beta_psi_star_prior,
    beta_psi_star_sd_prior = 1,
    eta_psi = 0.1,
    jj_psi = unit_species_revisit_summary |> dplyr::pull(unit_id),

    # theta-level inputs
    w_theta = w_theta,
    m_alpha = w_theta |> ncol(),
    w_theta_star = w_theta_star,
    m_alpha_star = w_theta_star |> ncol(),
    alpha_theta_star_prior = alpha_theta_star_prior,
    alpha_theta_star_sd_prior = 1,
    eta_theta = 0.1,
    jj_theta = unit_species_revisit_summary |> dplyr::pull(unit_id),

    sigma_theta_prior = 1.0,
    sigma_theta_prior_sd = 0.1,

    sigma_psi_prior = 1.0,
    sigma_psi_prior_sd = 0.1,

    # reduce_sum setting
    grainsize = 1
  )

```

The model may then be fit and the summary calculated:
```{r}
#| eval: false
n_chains <- 4
n_parallel_chains <- 4
n_threads_per_chain <- 4
n_refresh <- 0
n_warmup <- 3000
n_sample <- 3000

fit <- occ_2lvl_multispp_corr(stan_data,
                              n_chains = n_chains,
                              n_parallel_chains = n_parallel_chains,
                              n_threads_per_chain = n_threads_per_chain,
                              n_warmup = n_warmup,
                              n_sample = n_sample)

fit_summary <-
  fit$summary()
```

Optionally, load the saved example outputs included with the package:
```{r}
#| echo: true
fit_summary <-
  readRDS(system.file("./extdata/occupancy_2_multispecies_corr_fit_summary.rds",
                      package = "occstanhm"))

stan_data <-
  readRDS(system.file("./extdata/occupancy_2_multispecies_corr_stan_data.rds",
                      package = "occstanhm"))
```

And then examine the summary outputs and compare to simulated data:
```{r}
fit_summary |>
  filter(grepl("logit_psi", variable)) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(median)) |>
  print(n = Inf)

fit_summary |>
  filter(grepl("beta_psi\\[", variable)) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean)) |>
  print(n = Inf)

stan_data$unit_species_revisit_summary |>
  select(unit, species, psi_obs)

stan_data$dat_species |>
  mutate(psi_logit = qlogis(psi_vec),
         theta_logit = qlogis(theta_vec))

fit_summary |>
  filter(grepl("logit_theta", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean)) |>
  print(n = Inf)

fit_summary |>
  filter(grepl("alpha_theta\\[", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean)) |>
  print(n = Inf)

stan_data$dat_species |>
  mutate(psi_logit = qlogis(psi_vec),
         theta_logit = qlogis(theta_vec)) |>
  select(unit, species, theta_logit, theta_vec)
```


## 3-level model

Similar to the two-level model, a three-level model can be coded in Stan:

```{r comment=''}
#| echo: false
system.file("stan_models/tutorial/occ_3lvl_multispp_corr.stan",
            package = "occstanhm") |>
  readLines() |>
  cat(sep = "\n")
```

Likewise, data may be simulated and then formatted for Stan:

```{r}
#| eval: false
set.seed(123)
n_units_sim <- 10 # max is 26 due to number of letters
n_species_sim <- 3

n_revisits_per_site_in <- rpois(n_units_sim, 40)
hist(n_revisits_per_site_in)

n_samples_per_revisits <- rpois(sum(n_revisits_per_site_in), 30) + 1L
hist(n_samples_per_revisits,
     breaks = seq(0, 50, by = 1))

## species unit
dat_species <-
  expand_grid(unit = rep(letters[seq(1, n_units_sim)]),
              species = paste("species", seq(1, n_species_sim))) |>
  mutate(psi_vec = rbeta(n(), 3, 3),
         theta_vec = rbeta(n(), 3, 3),
         p_vec = rbeta(n(), 3, 3),
         k_subsamples = 6)

dat_species |>
  print(n = Inf)

## unit and revisit data
dat_revisit <-
  tibble(unit = rep(letters[seq(1, n_units_sim)],
                    times = n_revisits_per_site_in)) |>
  rowid_to_column("index") |>
  group_by(unit) |>
  mutate(revisit_id = dplyr::cur_group_rows() -
           min(dplyr::cur_group_rows()) + 1L) |>
  ungroup() |>
  select(unit, revisit_id)

dat_revisit |>
  print(n = Inf)

# sample
dat_sample <-
  tibble(
    unit = rep(dat_revisit |> pull(unit),
               times = n_samples_per_revisits),
    revisit_id = rep(dat_revisit |> pull(revisit_id),
                     times = n_samples_per_revisits),
  ) |>
  rowid_to_column("index") |>
  group_by(unit) |>
  mutate(sample_id = dplyr::cur_group_rows() -
           min(dplyr::cur_group_rows()) + 1L) |>
  select(-index) |>
  ungroup()

# merge together
dat <-
  dat_species |>
  full_join(dat_revisit, by = "unit",
            relationship = "many-to-many") |>
  mutate(z = rbinom(n(), 1, psi_vec)) |>
  full_join(dat_sample, by = c("unit", "revisit_id"),
            relationship = "many-to-many") |>
  arrange(unit, species) |>
  rowid_to_column("index") |>
  mutate(a = z * rbinom(n(), 1, theta_vec),
         y = a * rbinom(n(), k_subsamples, p_vec),
         a_obs = ifelse(y > 0, 1, 0)) |>
  ungroup() |>
  group_by(unit, revisit_id, species) |>
  mutate(z_obs = ifelse(sum(a_obs) > 0, 1, 0)) |>
  ungroup()

## Need to make sure arrange by unit first
## then within unit
unit_species_summary <-
  dat |>
  group_by(unit, species) |>
  summarize(psi_obs = mean(z_obs),
            unit_species_dat_start = min(index),
            unit_species_dat_stop = max(index),
            n_revisits_per_site = n(),
            .groups = "keep") |>
  mutate(logit_psi = qlogis(psi_obs))

unit_species_summary |>
  print(n = Inf)

revisit_species_summary <-
  dat |>
  group_by(unit, species, revisit_id) |>
  summarize(n_samples_per_revisit = n(),
            revisit_start = min(index),
            revisit_stop = max(index),
            z_obs = mean(z_obs),
            .groups = "drop") |>
  group_by(unit, species) |>
  mutate(revisit_start_in_unit_species =
         revisit_start - min(revisit_start) + 1L,
         revisit_stop_in_unit_species = revisit_stop - min(revisit_start) + 1L)

unit_species_revisit_summary <-
  revisit_species_summary |>
  rowid_to_column("index") |>
  group_by(unit, species) |>
  summarize(unit_species_revisit_start = min(index),
            unit_species_revisit_stop = max(index),
            n_revisits_per_unit_species = n(),
            psi_obs = mean(z_obs),
            .groups = "drop") |>
  mutate(unit_id = as.integer(factor(unit)))

## Setup parameters for model

# psi-level
x_psi <- model.matrix(~ species - 1, unit_species_summary)
x_psi |> head()
x_psi |> dim()

unit_species_summary_hyper <-
  unit_species_summary |>
  group_by(unit) |>
  summarize(n = n(),
            hyper_psi = mean(psi_obs),
            hyper_logit = mean(logit_psi))

x_psi_star <-
  model.matrix(~ 1, unit_species_summary_hyper)

beta_psi_star_prior <-
  matrix(0,
         nrow = x_psi_star |> ncol(),
         ncol = x_psi |> ncol())

# theta-level
w_theta <- model.matrix(~ species - 1, unit_species_summary)
w_theta
w_theta |> head()
w_theta |> dim()

w_theta_star <-
  model.matrix(~ 1, unit_species_summary_hyper)

alpha_theta_star_prior <-
  matrix(0,
         nrow = w_theta_star |> ncol(),
         ncol = w_theta |> ncol())

# p-level
v_p <- model.matrix(~ species - 1, unit_species_summary)
v_p
v_p |> head()
v_p |> dim()

v_p_star <-
  model.matrix(~ 1, unit_species_summary_hyper)

delta_p_star_prior <-
  matrix(0,
         nrow = v_p_star |> ncol(),
         ncol = v_p |> ncol())

## Save data in list for stan
stan_data <-
  # observation and latent observation data
  list(
    y = dat |> pull(y),
    k_subsamples = dat |> pull(k_subsamples),
    a_obs = dat |> pull(a),
    z_obs = revisit_species_summary |> pull(z_obs),
    # indexing
    unit_species_dat_start = unit_species_summary |>
      pull(unit_species_dat_start),
    unit_species_dat_stop = unit_species_summary |>
      pull(unit_species_dat_stop),
    revisit_start_in_unit = revisit_species_summary |>
      pull(revisit_start_in_unit_species),
    revisit_stop_in_unit = revisit_species_summary |>
      pull(revisit_stop_in_unit_species),
    unit_species_revisit_start = unit_species_revisit_summary |>
      pull(unit_species_revisit_start),
    unit_species_revisit_stop = unit_species_revisit_summary |>
      pull(unit_species_revisit_stop),
    # all different dimensions, such as "n"
    n_units = dat |> distinct(unit) |> nrow(),
    n_units_species = unit_species_summary |> nrow(),
    n_total_samples = dat |> nrow(),
    n_revisits_per_unit_species = unit_species_revisit_summary |>
      pull(n_revisits_per_unit_species),
    n_samples_per_revisit_species = revisit_species_summary |>
      pull(n_samples_per_revisit),
    n_total_revisits_species = revisit_species_summary |> nrow(),
    # psi-level inputs
    x_psi = x_psi,
    m_beta = x_psi |> ncol(),
    x_psi_star = x_psi_star,
    m_beta_star = x_psi_star |> ncol(),
    beta_psi_star_prior = beta_psi_star_prior,
    beta_psi_star_sd_prior = 1,
    eta_psi = 0.1,
    jj_psi = unit_species_revisit_summary |> dplyr::pull(unit_id),
    # theta-level inputs
    w_theta = w_theta,
    m_alpha = w_theta |> ncol(),
    w_theta_star = w_theta_star,
    m_alpha_star = w_theta_star |> ncol(),
    alpha_theta_star_prior = alpha_theta_star_prior,
    alpha_theta_star_sd_prior = 1,
    eta_theta = 0.1,
    jj_theta = unit_species_revisit_summary |> dplyr::pull(unit_id),
    # theta-level inputs
    v_p = v_p,
    m_delta = v_p |> ncol(),
    v_p_star = v_p_star,
    m_delta_star = v_p_star |> ncol(),
    delta_p_star_prior = delta_p_star_prior,
    delta_p_star_sd_prior = 1,
    eta_p = 0.1,
    jj_p = unit_species_revisit_summary |> dplyr::pull(unit_id),

    sigma_psi_prior = 1.0,
    sigma_psi_prior_sd = 0.1,

    sigma_theta_prior = 1.0,
    sigma_theta_prior_sd = 0.1,

    sigma_p_prior = 1.0,
    sigma_p_prior_sd = 0.1,
    # reduce_sum setting
    grainsize = 1
  )
```

The model may then be fit with Stan:

```{r}
#| eval: false
n_chains <- 4
n_parallel_chains <- 4
n_threads_per_chain <- 4
n_refresh <- 100
n_warmup <- 3000
n_sample <- 3000

fit <- occ_3lvl_multispp_corr(stan_data,
                              n_chains = n_chains,
                              n_parallel_chains = n_parallel_chains,
                              n_threads_per_chain = n_threads_per_chain,
                              n_refresh = n_refresh,
                              n_warmup = n_warmup,
                              n_sample = n_sample)
```

The saved results may optionally be loaded:

```{r}
#| echo: true
fit_summary <- system.file("extdata/occupancy_3_multispecies_corr_fit_summary.rds",
                   package = "occstanhm") |>
  readRDS()
```

And then the results examined:

```{r}
#| eval: false
fit_summary <-
  fit$summary()

fit_summary

fit_summary |>
  filter(grepl("logit_psi", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(median)) |>
  print(n = Inf)

fit_summary |>
  filter(grepl("beta_psi\\[", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean)) |>
  cbind(unit_species_revisit_summary |>
          arrange(species, unit) |>
          select(unit, species, psi_obs)) |>
  as_tibble() |>
  mutate(diff = prob - psi_obs) |>
  select(variable, unit, species, mean, median, prob, psi_obs, diff)

fit_summary |>
  filter(grepl("logit_theta", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean)) |>
  print(n = Inf)

fit_summary |>
  filter(grepl("alpha_theta\\[", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean)) |>
  cbind(dat_species |>
          arrange(species, unit) |>
          mutate(theta_logit = qlogis(theta_vec)) |>
          select(unit, species, theta_logit, theta_vec)) |>
  as_tibble() |>
  select(variable, unit, species, mean, median, theta_logit, prob, theta_vec)

fit_summary |>
  filter(grepl("logit_p\\[", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean)) |>
  print(n = Inf)

fit_summary |>
  filter(grepl("delta_p\\[", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean)) |>
  cbind(dat_species |>
          arrange(species, unit) |>
          mutate(p_logit = qlogis(p_vec)) |>
          select(unit, species, p_logit, p_vec)) |>
  as_tibble() |>
  select(variable, unit, species, mean, median, p_logit, prob, p_vec)

fit_summary |>
  filter(grepl("Omega_psi\\[", variable)) |>
  select(variable, mean, median) |>
  mutate(mean = round(mean, 2),
         median = round(median, 2),
         prob = plogis(mean))
```


## References

Erickson RA, Stich DS, Hebert JL. 2020. fishStan: Hierarchical Bayesian models for fisheries. U.S. Geological Survey software release. Reston, Va.
https://doi.org/10.5066/P9TT3ILO.

Erickson RA, Stich DS, Hebert JS. 2022. fishStan: hierarchical Bayesian models for fisheries. _Journal of Open Source Software_ 7.71: 3444.
