---
title: "Comparison to Tobler et al. (2019)"
author: "Richara A. Erickson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette compares the `occstanhm_2_fast` model to the Latent Variable model in Tobler et al. (2019; https://doi.org/10.1002/ecy.2754).
The Tobler et al. (2019; https://doi.org/10.1002/ecy.2754) code is not run and requires JAGS and supporting packages to be installed.

This example model also shows how to add non-hierarchical coefficients to a model in Stan.

```{r, echo=FALSE}
system.file("./compare_Tobler/compare_and_save.R", package = "occstanhm") |>
  knitr::read_chunk()
```

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

## Simulate  data

First, the package are loaded:

```{r used_packages}
```

Second, the Stan model settings are set:

```{r run_stan_model}
```

Next, the data is simulated and formatted for Stan:

```{r simulate_data}
```

The predictors for Stan may be checked for the $\psi$-level of the model:

```{r check_psi}
```

The detection (_p_) level inputs may also be formulated and checked:

```{r p_coef}
```

## Run occstanhm model

Now, the data may be formatted for Stan:

```{r save_as_stan}
```

Then, the model run in Stan:

```{r stan_run}
```

The summary may also be examined:

```{r fit_summary}
```

## Run Tobler mdoel

The data needs to be reformatted for the Tolber model:

```{r setup_tobler}
```

And then the Tobler et al. model run and post-processing done for both outputs:

```{r run_tobler}
```

## Lastly, the outputs may be compared and saved

```{r compare_outputs}
```



```
